---
title: "Applied Bayesian Modeling - module 6"
author: "Leontine Alkema"
date: "September 17, 2022"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(brms)
library(posterior)
library(tidybayes)
```

# Radon data 

The radon_data.Rmd file contains the code to read in the data and process it. We just work with the clean data here 

```{r}
dat0 <- read_csv(here::here("data/radon.csv"))
```

data for regression
```{r}
dat <- dat0 %>%
  mutate(y = log(activity), x_ura = log(ura_county) - mean(log(ura_county)))
```


# Model fitting using lm and brms

## frequentist/tradtional

simple fit 
```{r}
fit_lm <-  lm(y ~ 1, data = dat) 
summary(fit_lm)
```

here the same model fitting, just coded up differently in a pipe, with output in nice form
```{r}
dat %>%
  lm(y ~ 1, data = .) %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95) %>%
  select(-statistic, -p.value) #%>%
#  knitr::kable(format = "latex", digits = 2)

```


## Bayesian regression 
```{r, cache = TRUE}
fit <- brm(y ~ 1, data = dat,
        chains = 4, iter = 1000, warmup = 500, cores = getOption("mc.cores", 4))
```


Quick summary overview
```{r}
summary(fit) 
```

You can also pull out whatever you're interested in using the summarise_draws function, e.g.
```{r}
draws <- as_draws_array(fit, variable = "Intercept") # this is our mu
summarise_draws(draws, default_summary_measures())
```

To get a 95% CI:
```{r}
summarise_draws(draws, ~quantile(.x, probs = c(0.025, 0.975)))
```

These overview plots show, for each parameter, its histogram with posterior samples and its trace plot.
```{r}
plot(fit, variable = c("Intercept", "sigma"))
```

# Plots to show posterior samples

Joint density, using bins
```{r}
as_draws_df(fit, pars = c("b_Intercept", "sigma"))%>%
  ggplot(aes(x = sigma, y = b_Intercept)) +
  geom_bin2d() +
  theme_bw()

```

Joint density, estimated
```{r}
as_draws_df(fit, pars = c("b_Intercept", "sigma"))%>%
  ggplot(aes(x = sigma, y = b_Intercept)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")+
  theme_bw()

```


marginal densities
```{r}
as_draws_df(fit, pars = c("b_Intercept"))%>%
  ggplot(aes(x = b_Intercept)) +
  geom_histogram() +
  theme_bw()
```

```{r}
as_draws_df(fit, pars = c("sigma"))%>%
  ggplot(aes(x = sigma)) +
  geom_histogram() +
  theme_bw()
```



# MCMC diagnostics 
Traceplots and posterior densities. Note that you can find the help function with 
```{r}
#?plot.brmsfit
```

```{r}
plot(fit, variable = c("b_Intercept", "sigma")) 
```

Check Rhat and effective sample size
```{r}
summary(fit) 
#names(summary(fit))
summary(fit)$fixed
summary(fit)$spec_pars
```



# MCMC diagnostics in a less ideal setting...

Please note: These fits are based on settings that are NOT recommended. We are just creating an example here of a fit where the MCMC diagnostics (rightly) show that there are issues with the sampling. 
```{r, cache = TRUE}
fit_bad <- brm(y ~ 1, data = dat,
        chains = 4, iter = 400, cores = getOption("mc.cores", 4),
        control = list(adapt_delta = 0.4, max_treedepth = 4)
        # these are NOT recommended options, trying to create problems here!
)
```


```{r}
plot(fit_bad,  variable = c("b_Intercept",  "sigma")) 
```

```{r}
summary(fit_bad)
```

We see that the effective sample size are too low. Rhat is no longer equal to 1 (although still less than 1.05). 


## comparison fit

Here we create a fit with the same number of iterations for comparison 
```{r, cache = TRUE}
fit2_short <- brm(y ~ 1, data = dat,
        chains = 4, iter = 400, cores = getOption("mc.cores", 4))
```


```{r}
plot(fit2_short,  variable = c("b_Intercept",  "sigma")) 
```

```{r}
summary(fit2_short)
```



